---
title: "vignette"
output: html_document
date: "2023-12-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Binary Classification Vignette

For the purpose of this vignette, we will use data from the National Institute of Diabetes and Digestive Kidney Disease.

Activity: ...
Open a new RScript and copy-paste the chunk below to load the data
```{r}

# load packages
library(readr)
library(vip)
library(naniar)
library(tidymodels)
library(ISLR)
library(ISLR2)
library(tidyverse)
library(glmnet)
library(modeldata)
library(ggthemes)
library(janitor)
library(kableExtra)
library(yardstick)
library(kknn)
library(corrplot)
library(themis)
library(dplyr)
library(ggplot2)
library(scales)
library(rpart.plot)
library(discrim)
library(klaR)
library(plotly)
library(ranger)
library(xgboost)
tidymodels_prefer()

# read data
db <- read_csv("data/diabetes.csv")

```


Data Partitioning
First we will do cross-validation and data splitting.  
We then partition the diabetes data into training and test sets.
```{r}
set.seed(3435)
db_split <- initial_split(db, prop = 0.80,
                               strata = "Outcome")
db_train <- training(db_split)
db_test <- testing(db_split)
db_fold <- vfold_cv(db_train,v=4)
```

```{r}
db_train$Outcome <- factor(db_train$Outcome)
db_recipe_demo <- recipe(Outcome ~ ., data = db_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_predictors()) %>% 
  step_upsample(Outcome, over_ratio = 0.5, skip = FALSE)

prep(db_recipe_demo) %>% bake(new_data = db_train) %>% 
  group_by(Outcome) %>% 
  summarise(count = n())
```



Binary Classification Algorithm #1 




Model Fitting
```{r}

```


Accuracy Measures
```{r}

```

Compute a panel of precision of precision, recall, and F1 score
```{r}

```


Binary Classification Algorithm #2






Binary Classification Algorithm #3



...


